Co-evolution of lexical meaning \& pragmatic use

### 2017-03-27 paper/paper-v3.tex
revision text

\subsection{Expressivity and learnability}

@@

While natural languages are pressured for both expressivity and learnability, these forces may
pull in opposite directions. Their opposition becomes particularly clear when considering the
extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). A language consisting of a single
form-meaning association is easy to learn but lacking in expressivity. Conversely, a language
that lexicalizes a distinct form for a large number of different meanings is highly expressive
but challenging to acquire.

@@

An elegant formal approach to capture the interaction between expressivity and learnability is
the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. In
its simplest, discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in a population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the update step, (ii) the fitness $f_i$ of each type $i$ before the update, and
(iii) the probability $Q_{ji}$ that an agent who wants to imitate, adopt, or learn the type of
an agent with type $j$ ends up acquiring type $i$:
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align}
The RMD consists of two components: fitness-based selection and transmission biases. This
becomes most transparent when we consider an equivalent formulation in terms of a step-wise
application of the discrete-time replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$
and its subsequent multiplication with a mutation matrix $Q$:
\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\sum_k x_k f_k}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
If the transmission matrix $Q$ is trivial in the sense that $Q_{ji}=1$ whenever $j=i$, the
dynamic reduces to the replicator dynamic. The replicator dynamic is a model of fitness-based
selection in which the relative frequency of type $i$ will increase with a gradient
proportional to its average fitness in the population. This dynamic is popular and
versatile because it can be derived from many abstract processes of biological and cultural
transmission and selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}, including conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. If fitness
$f_i$ is the same for all types $i$, the replicator step is the identity map
$ \left ( \text{RD}(\vec{x}) \right )_i = x_i$ and the dynamic reduces to a process of
iteration of the transmission bias encoded in $Q$. In this way, the process in
(\ref{eq:RMD_discrete}), equivalently (\ref{eq:RMD_discrete_recast}), contains a model of
iterated learning \citep{griffiths+kalish:2007}.

@@

\subsubsection{Fitness \& fitness-based selection based on expressivity}\label{sec:expressivity}

Under the replicator dynamic the proportion of type $i$ in a population will increase or
decrease as a function of its relative fitness $f_i$. In the context of language evolution,
fitness is usually associated with expressivity, i.e., the ability to successfully communicate
with other language users from the same population
\citep[e.g.,][]{nowak+krakauer:1999,nowak+etal:2000, nowak+etal:2002}. Under a biological
interpretation the assumption is that organisms have a higher chance of survival and
reproduction if they are able to share and receive useful information via communication with
peers. Under a cultural interpretation the picture is that agents themselves strive towards
communicative success and therefore occasionally adapt or revise their behavior to achieve
higher communicative success (see \citealt[\S3.3]{benz+etal:2005b} for discussion).

@@

Languages are shaped not only by functionalist forces towards greater expressivity. Another
important factor is the fidelity with which linguistic knowledge is transmitted. Among others,
linguistic production can be prone to errors, states or messages may be perceived incorrectly,
and multiple languages may be compatible with the data learners are exposed to. These sources
of uncertainty introduce variation in the transmission of linguistic knowledge from one
generation to the next. In particular, learning biases in the iterated transmission process can
influence language evolution substantially.

@@

Expressivity and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamic in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the step, its fitness $f_i$, and the probability that a learner infers
$i$ when observing the behavior of a type-$j$ agent. Fitness-based selection is here thought of
not as biological (fitness as expected relative number of offspring) but cultural (fitness as
likelihood of being imitated or repeated) evolution, since the types that the dynamic operates
on are pairs consisting of a lexicon and a pragmatic use pattern. A type's expressivity depends
on its communicative efficiency within a population while its learnability depends on the
fidelity by which it is inferred by new generations of learners. The learners' task is
consequently to perform a joint inference over types of linguistic behavior and lexical
meaning.

@@

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOTs) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008, piantadosi+etal:underreview,
  piantadosi+etal:2012a} and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which concepts can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
motivate and formalize a preference of learners for simpler semantic representations
\citep{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a,
  kirby+etal:2015,piantadosi+etal:underreview}. A concept's complexity is a function of its
derivation cost in a weighed generative LOT.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/fig1-onlyr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and majority types, when other
  than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a
  pressure for expressivity. The right-most plots zoom in on only the proportion of target types. \tb{change plot to maj. type}}
\label{fig:only-R}
\end{figure}

As expected, selection based on communicative success is sensitive to $\lambda$ as it
influences signaling behavior, which in turn determines communicative success. This is
showcased in Figure \ref{fig:only-R}, which depicts the proportion of target types in $5$
independent populations after $50$ replicator steps. The plot also shows the proportion of the
\emph{majority type}, i.e., the type with the highest proportion in the final population. With
low $\lambda$ many types have very similar behavior, so that evolutionary selection lacks grip
and becomes very slow. The result is a very long transition with near stagnancy in a rather
homogeneous population with many types. Conversely, higher $\lambda$ promotes more rational
linguistic behavior, widening the gap in expressivity between types and promoting more
homogeneous populations. As suggested by Figure \ref{fig:only-R}, the majority in most
populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure only for
expressivity does not lead to a prevalence of target types under any $\lambda$-value. For
instance, with $\lambda = 20$ $1000$ independent populations only had $11$ cases in which
the target type was the majority type, corresponding to a mean proportion of $0.003$ across
populations. By contrast, in $913$ cases the majority types had $\Lbound$ with close to an even
share between literal ($454$) and pragmatic types ($459$), corresponding to a mean proportion
of about $.48$ taken together.

@@

The effect of iterated learning without a pressure for expressivity using either posterior
sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in
Figure \ref{fig:only-M} together with the prior over types. The prior shows that while users of
$\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are
nevertheless more advantaged than others, such as $\Lbound$, in virtue of the relatively simple
semantics they conventionalize (see Section~\ref{sec:an-induct-learn}). Crucially, $\Llack$
enables its users to convey each state with a single message when combined with pragmatic
reasoning provided sufficiently high $\lambda$. This makes it less likely to be confused with
other types if the learning data is not too sparse ($k \geq 5$). Put differently, learners have
a higher propensity to infer pragmatic $\Llack$ when the teacher's type produces very similar
data, such as when using $\Lbound$. Moreover, $\Llack$ is also less likely to be confused with
types with different observable behavior because its pragmatic use approximates a one-to-one
form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior
increases their proportion in the population.

However, in contrast to a pressure only for expressivity with high $\lambda$ (see
Figure~\ref{fig:only-R}), learnability alone does not succeed in selecting for a single
prevalent type. Indeed, all six target types tend to coexist at roughly equal proportion. Each
is passed on to the next generation with the same faithfulness and, differently from a pressure
for communicative success, they do not stand in competition with each other. In $1000$
independent populations with $\lambda = 20$ all majority types were target types,
with each reaching approximately the same proportion of users in the population. As
with expressivity only, low values of $\lambda$ make the differences in observable behavior
across types less pronounced and therefore reflect the learners' inductive bias more
faithfully, favoring functionally deficient but a priori preferred types such as those that use
$\Lall$. A pressure for learnability alone may consequently lead to a spread of communicatively
suboptimal types that are easier to learn. In the extreme, when $l = 1$ and $\lambda = 1$ all
of $1000$ independent populations had users of $\Lall$ as majority types.

\subsubsection{Combining pressures of expressivity and learnability}

Pressures for communicative success or learnability are not sufficient on their own to have a
single target type dominate the population. When pressured for expressivity, the slight
communicative advantage of $\Lbound$ users leads to the its prevalence. When pressured for
learnability, pragmatic $\Llack$ is promoted over functionally similar but semantically more
complex alternatives such as $\Lbound$. However, learnability alone does not foment the
propagation of a single target type across the population.

@@

To recapitulate, other than the involvement of pressure on both expressivity and learnability,
the resulting proportion of pragmatic $\Llack$ speakers primarily hinges on three
factors. First, the degree, captured by $\lambda$, to which agents try to maximize
communicative success from their own subjective point of view. Second, the inductive bias,
which leads learners to prefer simpler over more complex semantic representations in
acquisition. Lastly, the learning behavior, captured by parameter $l$, where approximating a
MAP strategy magnifies the effects of the learning bias in tandem with replication.

In summary, target types, which represent the majority view of scalar implicatures, can come to
dominate the population if three assumptions are met: (i) language is pressured toward both
expressivity and learnability; (ii) pragmatic language use is an option; (iii) learners prefer
simpler over more complex lexical representations and exhibit a tendency towards the
acquisition of the type that best explains the learning data.

@@

The approach introduced here combines game theoretic models of functional pressure towards
efficient communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on
(iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener
types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012,
  franke+jaeger:2014} as well as reasoning about unobservable lexical representations
\citep{bergen+etal:2012,bergen+etal:2016}. This allows a conceptual investigation of the
co-evolution of conventional meaning and pragmatic language use. Main contributions of the
model are (i) its modular separation of expressivity and learnability on evolutionary
trajectories, (ii) the characterization of language learning as a joint inference over
pragmatic behavior and lexical meaning, and (iii) the possibility to trace the co-evolution of
conventional semantics and pragmatic use.

\mf{can this abstract be shortened (by a factor of $.4$)?} \tb{See notes at end} The interaction of expressivity and
learnability in language evolution has also recently been addressed by
\citet{kirby+etal:2015}. In contrast to our proposal, Kirby et al. model expressivity as
exerting its force only in the production of learning data. The degree of mutual understanding
of interlocutors central to replication and to our notion of expressivity is absent. That is,
while our proposal combines bidirectional horizontal transmission with its vertical and
unidirectional counterpart, Kirby et al. only consider the latter's influence. Our reasoning
behind the inclusion of the former lies in the empirical and theoretical observation that
learnability alone can lead to the selection of functionally defective languages, as showcased
by $\Lall$. This outcome has been reported in a number of laboratory experiments where subjects
learned and subsequently reproduced the language produced by a previous participant over
multiple iterations, leading to a proliferation of ambiguous languages that associate many
meanings with a single form (see, e.g., \citealt{silvey+etal:2014} and experiment 1 in
\citealt{kirby+etal:2008}). In contrast, experiments involving an interactive component have
been found to foster languages that enable interlocutors to communicatively distinguish
meanings more accurately (e.g., \citealt{fay+etal:2013}; for a review of laboratory results
under the iterated learning paradigm and further discussion see \citealt{kirby+etal:2015,
  tamariz+kirby:2016}). It is not evident how to compare these empirical findings given that
they consider distinct meaning spaces, modes of transmission, iterations and feedback given to
participants. However, on a general level they may suggest that there is an important
difference between a language generating learnable linguistic data and its actual performance
as a means of information transfer. That is, we contend that successful information transfer in
a linguistic community is central to the adoption of a communication system and that this
measure may not be adequately reflected by production alone.

@@

While the results of this case study are interesting, they also raise a number of critical
issues. First of all, while many favorable parameter settings exist which lead to a prevalence
of target types, other types are usually represented in non-negligible proportions. This may
just be a technical quirk of the mutator step. But there is a related issue of empirical
importance. Several experimental studies on scalar implicatures suggest that participants can
be classified as either semantic or pragmatic users of, in particular, \emph{some}
\citep[e.g.][]{BottNoveck2004:Some-Utterances,NieuwlandDitman2010:On-the-incremen,DegenTanenhaus2012:Processing-Scal}. The
former consistently accept \emph{some} where \emph{all} would be true as well, the latter do
not. Interestingly, in our simulations when a target type is the majority type an inflated
proportion of the population uses compatible lexica with a lexicalized upper bound. \mf{Thomas,
  please check this; I'm making it up; total conjecture; it might be true but I don't know for
  sure} In other words, we do find a similar co-existence of semantic and pragmatic
types. Whether this analogy has any further explanatory value is an interesting path for future
exploration.

@@

\section{Conclusion}
The cultural evolution of meaning is influenced by intertwined pressures. We set out to
investigate this process by putting forward a model that combines a pressure toward successful
information transfer with perturbations that may arise in the transmission of linguistic
knowledge in acquisition. Its objects of selection and replication are pairs of lexical
meanings and patterns of language use. This allows the model to trace the interaction between
conventional meaning and pragmatic use. Additionally, it takes the challenge serious of neither
semantics nor pragmatics being directly observable. Instead, learners need to infer these
unobservables from overt data that results from their combination.  These components and their
mutual influence were highlighted in a case study on the lack of lexical upper-bounds in weak
scalar expressions that showed that, when pressured for learnability and expressivity, the
former force drives for simpler semantic representations inasmuch as pragmatics can compensate
for lack of expressivity in use. That is, the relative learning advantage of simpler semantics
in combination with functional pressure in use may offer an answer to why natural languages
fail to lexicalize systematic pragmatic inferences. And, more broadly, to a division of labor
between semantics and pragmatics.

>>>>

\subsection{Communicative success and learnability}

@@

While natural languages are pressured for both communicative success and learnability, these forces may
pull in opposite directions. Their opposition becomes particularly clear when considering the
extreme (cf. \citealt{kemp+regier:2012,kirby+etal:2015}). A language consisting of a single
form-meaning association is easy to learn but lacking in expressivity. Conversely, a language
that lexicalizes a distinct form for a large number of different meanings is highly expressive
but challenging to acquire.

@@

An elegant formal approach to capture the interaction between communicative success and learnability is
the \emph{replicator mutator dynamic}
\citep{Hofbauer1985:The-Selection-M,nowak+etal:2000,NowakKomarova2001:Evolution-of-Un,hofbauer+sigmund:2003,Nowak2006:Evolutionary-Dy}. In
its simplest, discrete-time formulation, the RMD defines the frequency $x'_i$ of each type $i$
in a population at the next time step as a function of: (i) the frequency $x_i$ of each type
$i$ before the update step, (ii) the fitness $f_i$ of each type $i$ before the update, and
(iii) the probability $Q_{ji}$ that an agent who wants to imitate, adopt, or learn the type of
an agent with type $j$ ends up acquiring type $i$:
\begin{align}
  \label{eq:RMD_discrete}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align}
The RMD consists of two components: fitness-based selection and transmission biases. This
becomes most transparent when we consider an equivalent formulation in terms of a step-wise
application of the discrete-time replicator dynamic \citep{TaylorJonker1978:Evolutionary-St} on the initial population vector $\vec{x}$
and its subsequent multiplication with a mutation matrix $Q$:
\begin{align}
  \label{eq:RMD_discrete_recast}
  x'_i & = (\text{M}(\text{RD}(\vec{x})))_i\,,
\end{align}
where
\begin{align*}
      \left ( \text{RD}(\vec{x}) \right )_i 
         = \frac{x_i f_i}{\sum_k x_k f_k}
 \ \ \ \ \text{and} \ \ \ \ 
  (\text{M}(\vec{x}))_i = (\vec{x} \cdot Q)_i = \left ( \sum_j
          x_j Q_{ji} \right)_i\,.
\end{align*}
If the transmission matrix $Q$ is trivial in the sense that $Q_{ji}=1$ whenever $j=i$, the
dynamic reduces to the replicator dynamic. The replicator dynamic is a model of fitness-based
selection in which the relative frequency of type $i$ will increase with a gradient
proportional to its average fitness in the population. This dynamic is popular and
versatile because it can be derived from many abstract processes of biological and cultural
transmission and selection \citep[for overview and several derivations
see][]{Sandholm2010:Population-Game}, including conditional imitation
\citep[e.g.,][]{Helbing1996:A-Stochastic-Be,Schlag1998:Why-Imitate-and} or reinforcement
learning \citep[e.g.,][]{BorgersSarin997:Learning-Throug,Beggs2005:On-the-Converge}. If fitness
$f_i$ is the same for all types $i$, the replicator step is the identity map
$ \left ( \text{RD}(\vec{x}) \right )_i = x_i$ and the dynamic reduces to a process of
iteration of the transmission bias encoded in $Q$. In this way, the process in
(\ref{eq:RMD_discrete}), equivalently (\ref{eq:RMD_discrete_recast}), contains a model of
iterated learning \citep{griffiths+kalish:2007}.

@@

\subsubsection{Fitness \& fitness-based selection based on communicative success}\label{sec:expressivity}

Under the replicator dynamic the proportion of type $i$ in a population will increase or
decrease as a function of its relative fitness $f_i$. In the context of language evolution,
fitness is usually associated with the ability to successfully communicate
with other language users from the same population
\citep[e.g.,][]{nowak+krakauer:1999,nowak+etal:2000, nowak+etal:2002}. Under a biological
interpretation the assumption is that organisms have a higher chance of survival and
reproduction if they are able to share and receive useful information via communication with
peers. Under a cultural interpretation the picture is that agents themselves strive towards
communicative success and therefore occasionally adapt or revise their behavior to achieve
higher communicative success (see \citealt[\S3.3]{benz+etal:2005b} for discussion).

@@

Languages are shaped not only by functionalist forces towards greater communicative success. Another
important factor is the fidelity with which linguistic knowledge is transmitted. Among others,
linguistic production can be prone to errors, states or messages may be perceived incorrectly,
and multiple languages may be compatible with the data learners are exposed to. These sources
of uncertainty introduce variation in the transmission of linguistic knowledge from one
generation to the next. In particular, learning biases in the iterated transmission process can
influence language evolution substantially.

@@

Communicative success and learnability are central to the cultural evolution of language. These
components can be modelled, respectively, as replication based on a measure of fitness in terms
of communicative efficiency and iterated Bayesian learning. Their interaction is described by
the discrete time replicator mutator dynamic in (\ref{eq:RMD_discrete}), repeated here:
\begin{align*}
  x'_i = \sum_j Q_{ji} \frac{x_jf_j}{\sum_k x_k f_k}\,.
\end{align*}
This equation defines the frequency $x'_i$ of type $i$ at the next time step, based on its
frequency $x_i$ before the step, its fitness $f_i$, and the probability that a learner infers
$i$ when observing the behavior of a type-$j$ agent. Fitness-based selection is here thought of
not as biological (fitness as expected relative number of offspring) but cultural (fitness as
likelihood of being imitated or repeated) evolution, since the types that the dynamic operates
on are pairs consisting of a lexicon and a pragmatic use pattern. A type's communicative success depends
on how well it communicates within a population while its learnability depends on the
fidelity by which it is inferred by new generations of learners. The learners' task is
consequently to perform a joint inference over types of linguistic behavior and lexical
meaning.

@@

There is a growing effort to develop empirically testable representational languages that allow
for the measure of semantic complexity. For instance, so-called {\em languages of thought}
(LOTs) have been put to test in various rational probabilistic models that show encouraging
results (see, e.g., \citealt{katz+etal:2008, piantadosi+etal:underreview,
  piantadosi+etal:2012a} and \citealt{piantadosi+jacobs:2016} for recent discussion). At its
core, a LOT defines a set of operations and composition rules from which concepts can be
derived. As a first approximation and for the sake of concreteness, we follow this approach to
motivate and formalize a preference of learners for simpler semantic representations
(\citealt{feldman:2000, chater+vitanyi:2003, piantadosi+etal:2012a,kirby+etal:2015}, \citeauthor{piantadosi+etal:underreview} under review). A concept's complexity is a function of its
derivation cost in a weighed generative LOT.

@@

\begin{figure}[t]
\centering
\includegraphics[width=\textwidth,height=8cm, keepaspectratio]{./plots/fig1-onlyr}
\caption{Stacked proportion of pragmatic $\Llack$-style types and majority types, when other
  than pragmatic $\Llack$, in $5$ independent populations after $50$ generations under only a
  pressure for communicative success. The right-most plots zoom in on only the proportion of target types. \tb{change plot to maj. type}}
\label{fig:only-R}
\end{figure}

As expected, selection based on communicative success is sensitive to $\lambda$ as it
influences signaling behavior, which in turn determines communicative success. This is
showcased in Figure \ref{fig:only-R}, which depicts the proportion of target types in $5$
independent populations after $50$ replicator steps. The plot also shows the proportion of the
\emph{majority type}, i.e., the type with the highest proportion in the final population. With
low $\lambda$ many types have very similar behavior, so that evolutionary selection lacks grip
and becomes very slow. The result is a very long transition with near stagnancy in a rather
homogeneous population with many types. Conversely, higher $\lambda$ promotes more rational
linguistic behavior, widening the gap in expressivity between types and promoting more
homogeneous populations. As suggested by Figure \ref{fig:only-R}, the majority in most
populations is not one of the six pragmatic $\Llack$-style types. That is, a pressure only for
communicative success does not lead to a prevalence of target types under any $\lambda$-value. For
instance, with $\lambda = 20$ $1000$ independent populations only had $11$ cases in which
the target type was the majority type, corresponding to a mean proportion of $0.003$ across
populations. By contrast, in $913$ cases the majority types had $\Lbound$ with close to an even
share between literal ($454$) and pragmatic types ($459$), corresponding to a mean proportion
of about $.48$ taken together.

@@

The effect of iterated learning without a pressure for communicative success using either posterior
sampling ($l = 1$) or a stronger tendency towards posterior maximization ($l = 15$) is shown in
Figure \ref{fig:only-M} together with the prior over types. The prior shows that while users of
$\Llack$ are not the most favored by the inductive bias (compared, e.g., to $\Lall$) they are
nevertheless more advantaged than others, such as $\Lbound$, in virtue of the relatively simple
semantics they conventionalize (see Section~\ref{sec:an-induct-learn}). Crucially, $\Llack$
enables its users to convey each state with a single message when combined with pragmatic
reasoning provided sufficiently high $\lambda$. This makes it less likely to be confused with
other types if the learning data is not too sparse ($k \geq 5$). Put differently, learners have
a higher propensity to infer pragmatic $\Llack$ when the teacher's type produces very similar
data, such as when using $\Lbound$. Moreover, $\Llack$ is also less likely to be confused with
types with different observable behavior because its pragmatic use approximates a one-to-one
form-meaning mapping. As a consequence, a stronger propensity to maximize the posterior
increases their proportion in the population.

However, in contrast to a pressure only for communicative success with high $\lambda$ (see
Figure~\ref{fig:only-R}), learnability alone does not succeed in selecting for a single
prevalent type. Indeed, all six target types tend to coexist at roughly equal proportion. Each
is passed on to the next generation with the same faithfulness and, differently from a pressure
for communicative success, they do not stand in competition with each other. In $1000$
independent populations with $\lambda = 20$ all majority types were target types,
with each reaching approximately the same proportion of users in the population. As
with a pressure only for communicative success, low values of $\lambda$ make the differences in observable behavior
across types less pronounced and therefore reflect the learners' inductive bias more
faithfully, favoring functionally deficient but a priori preferred types such as those that use
$\Lall$. A pressure for learnability alone may consequently lead to a spread of communicatively
suboptimal types that are easier to learn. In the extreme, when $l = 1$ and $\lambda = 1$ all
of $1000$ independent populations had users of $\Lall$ as majority types.

\subsubsection{Combining pressures of communicative success and learnability}

Pressures for communicative success or learnability are not sufficient on their own to have a
single target type dominate the population. When pressured for communicative success, the slight
communicative advantage of $\Lbound$ users leads to the its prevalence. When pressured for
learnability, pragmatic $\Llack$ is promoted over functionally similar but semantically more
complex alternatives such as $\Lbound$. However, learnability alone does not foment the
propagation of a single target type across the population.

@@

To recapitulate, other than the involvement of pressure on both communicative success and learnability,
the resulting proportion of pragmatic $\Llack$ speakers primarily hinges on three
factors. First, the degree, captured by $\lambda$, to which agents try to maximize
communicative success from their own subjective point of view. Second, the inductive bias,
which leads learners to prefer simpler over more complex semantic representations in
acquisition. Lastly, the learning behavior, captured by parameter $l$, where approximating a
MAP strategy magnifies the effects of the learning bias in tandem with replication.

In summary, target types, which represent the majority view of scalar implicatures, can come to
dominate the population if three assumptions are met: (i) language is pressured toward both
communicative success and learnability; (ii) pragmatic language use is an option; (iii) learners prefer
simpler over more complex lexical representations and exhibit a tendency towards the
acquisition of the type that best explains the learning data.

@@

The approach introduced here combines game theoretic models of functional pressure towards
efficient communication \citep{nowak+krakauer:1999}, effects of transmission perturbations on
(iterated) language learning \citep{griffiths+kalish:2007}, probabilistic speaker and listener
types of varied degrees of pragmatic sophistication \citep{frank+goodman:2012,
  franke+jaeger:2014} as well as reasoning about unobservable lexical representations
\citep{bergen+etal:2012,bergen+etal:2016}. This allows a conceptual investigation of the
co-evolution of conventional meaning and pragmatic language use. Main contributions of the
model are (i) its modular separation of communicative success and learnability on evolutionary
trajectories, (ii) the characterization of language learning as a joint inference over
pragmatic behavior and lexical meaning, and (iii) the possibility to trace the co-evolution of
conventional semantics and pragmatic use.

With respect to (i), \citet{kirby+etal:2015} propose a comparable model of the interaction
between a lexicon's expressivity and its learnability. One of its main differences to our proposal lies in 
the former force exerting pressure only in the production of learning data. The pressure towards communicative success 
within a population central to replication is absent. Put differently, we consider a pressure for mutual understanding 
that may indirectly select for expressive types, i.e., those that can convey states unequivocally, whereas Kirby et al. only
consider the bearing that the latter ability has on the production of learnable data. We see two main reasons
for considering communicative success rather than expressivity, and to only considering it in learning. First, learnability alone can 
lead to the acquisition of functionally defective languages, as showcased by $\Lall$ (cf. \citealt{kirby+etal:2008,silvey+etal:2014}, and \citealt{fay+etal:2013} for review of laboratory results). Second and more importantly, types may be equally expressive but their performance as means of information transfer crucially depends not only on themselves but on the population they find themselves in (cf. the competition of types of a kind in Figure \ref{fig:rmd} and their lack of competition in Figure \ref{fig:only-M}). That is, we contend that having an expressive lexicon that generates learnable data does not  in itself capture a type's arguably central communicative function of transferring information to peers. Under this view, a type's expressiveness is secondary to its ability to communicative successfully. Acquiring a maximally expressive type is of little use when it is not understood in its population.

@@

While the results of this case study are interesting, they also raise a number of critical
issues. First of all, while many favorable parameter settings exist which lead to a prevalence
of target types, other types are usually represented in non-negligible proportions. This may
just be a technical quirk of the mutator step. But there is a related issue of empirical
importance. Several experimental studies on scalar implicatures suggest that participants can
be classified as either semantic or pragmatic users of, in particular, \emph{some}
\citep[e.g.][]{BottNoveck2004:Some-Utterances,NieuwlandDitman2010:On-the-incremen,DegenTanenhaus2012:Processing-Scal}. The
former consistently accept \emph{some} where \emph{all} would be true as well, the latter do
not. Interestingly, in our simulations when a target type is the majority type an inflated
proportion of the population uses compatible lexica with a lexicalized upper bound. Particularly in those parameter
settings where the prior influences the outcome less. In other words, we do find a tendency toward a similar co-existence
of semantic and pragmatic types. Whether this analogy has any further explanatory value is an interesting path for future
exploration. \tb{I some qualifications. While it is indeed true that $\Lbound$'s proportion is inflated relative to other types in the hypothesis space, it is never the second most prolific type (which is what the above hopefully is not taken to suggest, otherwise it should be changed). As is to be expected, $\Lbound$ fares best when $l$ is low and $\lambda$ is high. Even then, it usually ranks around $5$-th with a mean proportion of its kind of about $0.01 - 0.05$. As $l$ increases, it gets pushed back. Second or third place are usually instead literal $\Llack$ or types that lexicalize ``all'', ``not all'' and ``none''. In short; yes, there is a inflated proportion but only compared to the whole population. Other particular types are usually more prolific. This is very likely due to the strong prior disadvantage of $\Llack$. In light of this I'm uncertain whether we want to keep this paragraph as is. I tend towards `yes' but let me know if you think otherwise.}

@@

\section{Conclusion}
The cultural evolution of meaning is influenced by intertwined pressures. We set out to
investigate this process by putting forward a model that combines a pressure toward successful
information transfer with perturbations that may arise in the transmission of linguistic
knowledge in acquisition. Its objects of selection and replication are pairs of lexical
meanings and patterns of language use. This allows the model to trace the interaction between
conventional meaning and pragmatic use. Additionally, it takes the challenge serious of neither
semantics nor pragmatics being directly observable. Instead, learners need to infer these
unobservables from overt data that results from their combination.  These components and their
mutual influence were highlighted in a case study on the lack of lexical upper-bounds in weak
scalar expressions that showed that, when pressured for learnability and communicative success, the
former force drives for simpler semantic representations inasmuch as pragmatics can compensate
for lack of expressivity in use. That is, the relative learning advantage of simpler semantics
in combination with functional pressure in use may offer an answer to why natural languages
fail to lexicalize systematic pragmatic inferences. And, more broadly, to a division of labor
between semantics and pragmatics.
